---
title: "Machine Learning: <br><small><b>Peer Assessment - Course Project</b></small>"
author: "<a href='mailto:svicente99@yahoo.com' title='twitter:@svicente99'>Sergio Vicente</a>"
twitter: "@svicente99"
date: "May, 2015"
output: 
  html_document:
    keep_md: true
transition: fade
transition-speed: fast
subtitle: "<span style='color:#5882FA;font-size:0.8em'>Human Activity Monitoring</span>" 
---

<br><br>
You may get source code of this at <https://github.com/svicente99/MachLearn_PeerAssessment> 

And html version is available at RPubs: <http://rpubs.com/svicente99/MachLearn_Peer_Assesment2>

* * * *

### Synopsis

------------

#### Introduction


---------

#### Data Sourcing

The data for this assignment can be downloaded from the course web site:

* Dataset: <a href="http://groupware.les.inf.puc-rio.br/har.">Activity Data</a> [__ MBytes]

There is also some documentation of the database available. See above links to know how some of the variables are constructed/defined.

---------

### Data Processing

Setting parameters and files to be processed:

```{r, cache=TRUE}

# MAIN PARAMETERS
DATA_FOLDER <- "./data"  # subdirectory of current named 'data'
URL_DATA <- "https://d396qusza40orc.cloudfront.net/predmachlearn"
CSV_TRAIN <- "pml-training.csv"
CSV_TEST <- "pml-testing.csv"
# -----------------------------------------------------------------

# cache file used in this job
cached_file1 = paste( DATA_FOLDER, CSV_TRAIN, sep = "/" );
cached_file2 = paste( DATA_FOLDER, CSV_TEST, sep = "/" );

```

Getting CSV data files and assembling the main datas frame (df_*):

```{r,cache=TRUE,echo=FALSE}

read_csv <- function(cached_file) 
{
  # check if the csv files are already saved ("cached") onto disk
  
  if( !file.exists(cached_file) ) {
      url = paste(URL_DATA, ZIP_FILE, sep="")
      if( !file.exists(DATA_FOLDER) )  dir.create(DATA_FOLDER)
  	  if( !file.exists(cached_file) ) { 
  		  # if it is not available --> Download it !
  		  downloadStatus = download.file(url, method="internal", destfile=cached_file)
  	      if(downloadStatus != 0)  stop("Download failed from URL: ", url)
  	  }
  } 
  # read data from zip file already writen on disk
  df <- read.csv(cached_file)
  return(df)
}

df_train <- read_csv(cached_file1)
df_test <- read_csv(cached_file2)

# above, we have a extract of ten lines of main data available to analyse
colnames(df_train)
print("Dimension of training data - rows x cols")
dim(df_train)
print("Dimension of test data - rows x cols")
dim(df_test)
```

```{r,cache=TRUE,echo=FALSE}
head(df_train["classe",])
```

* * * *

## <u>Conclusion</u>



* * * *
